---
title: "IST 707 Final Project - Student Educational Outcomes"
author: "Christoper Murphy, Padmaja Kurumaddali, La Monte Yarroll"
date: "Septemner 07, 2024"
output: word_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include=FALSE}
#first, install and load required libraries
#using function provided by Professor Gary Krudys.

#specify the packages needed
packages=c("tidyverse", "ggplot2", "mclust", "rpart", "rpart.plot", "caret", "arules", "arulesviz")

#use this function to check if each package is on the local machine
#if a package is installed, it will be loaded
#if any are not, the missing package(s) will be installed and loaded
package.check <- lapply(packages, FUN = function(x) {
  if (!require(x, character.only = TRUE)) {
    install.packages(x, dependencies = TRUE)
    library(x, character.only = TRUE)
  }
})

#verify they are loaded
search()
```


```{r eval=FALSE, include=FALSE}
#References used to inform code choices:

```


# Introduction
 

# Analysis and Models


## About the data


### Dataset Summary
The following dataset was located on Kaggle at the following page:

https://www.kaggle.com/code/mdismielhossenabir/student-s-dropout-and-academic-success

The original source at UCI is here:

https://archive.ics.uci.edu/dataset/697/predict+students+dropout+and+academic+success

This dataset contains information on students' marital status, application details, previous qualifications, family background, admission grades, financial status, and academic performance. The target variable indicates whether a student dropped out or achieved academic success. The data spans multiple semesters and includes economic indicators such as unemployment rate, inflation rate, and GDP. Researchers can leverage this dataset to develop models for predicting dropout and academic success, contributing to the understanding of factors influencing educational outcomes.


### Data Retrieval
Data structure structure when read in from the .CSV file:

```{r echo=FALSE}
filePath = "studentdata.csv"
studentSuccessData = read.csv2(filePath, stringsAsFactors = FALSE, dec=".")
str(studentSuccessData)
```


#### Initial Data Sample
A sample set of records:
```{r echo=FALSE}
head(studentSuccessData, 5)
```

#### Check for Missing Values

```{r echo=FALSE}
colSums(is.na(studentSuccessData))
```



\pagebreak
### Data Attribute Summary

The Kaggle data card can also be found on the following page
https://www.kaggle.com/code/mdismielhossenabir/student-s-dropout-and-academic-success/input

### Attribute Summary:
* Marital Status: The marital status of the student
* Application mode: The mode through which the student applied for admission
* Application order: The order in which the student applied for admission.
* Course: The academic course or program the student is enrolled in.
* Daytime/evening attendance: The attendance mode of the student (e.g., Daytime, Evening).
* Previous qualification: The highest academic qualification the student attained before the current enrollment.
* Previous qualification (grade): The grade or score achieved in the previous qualification.
* Nacionality: The nationality of the student.
* Mother's qualification: The highest academic qualification of the student's mother.
* Father's qualification: The highest academic qualification of the student's father.
* Mother's occupation: The occupation of the student's mother.
* Father's occupation: The occupation of the student's father.
* Admission grade: The grade or score obtained during the admission process.
* Displaced: Indicates whether the student has been displaced from their previous location.
* Educational special needs: Flags if the student has any educational special needs.
* Debtor: Indicates if the student is a debtor.
* Tuition fees up to date: Indicates whether the student's tuition fees are up to date.
* Gender: The gender of the student.
* Scholarship holder: Flags if the student is a holder of a scholarship.
* Age at enrollment: The age of the student at the time of enrollment.
* International: Indicates if the student is an international student.
* Curricular units 1st sem (credited): The number of curricular units credited in the first semester.
* Curricular units 1st sem (enrolled): The number of curricular units enrolled in the first semester.
* Curricular units 1st sem (evaluations): The number of evaluations conducted in the first semester.
* Curricular units 1st sem (approved): The number of curricular units approved in the first semester.
* Curricular units 1st sem (grade): The overall grade for curricular units in the first semester.
* Curricular units 1st sem (without evaluations): Curricular units in the first semester without evaluations.
* Curricular units 2nd sem (credited): The number of curricular units credited in the second semester.
* Curricular units 2nd sem (enrolled): The number of curricular units enrolled in the second semester.
* Curricular units 2nd sem (evaluations): The number of evaluations conducted in the second semester.
* Curricular units 2nd sem (approved): The number of curricular units approved in the second semester.
* Curricular units 2nd sem (grade): The overall grade for curricular units in the second semester.
* Curricular units 2nd sem (without evaluations): Curricular units in the second semester without evaluations.
* Unemployment rate: The unemployment rate at the time of data collection.
* Inflation rate: The inflation rate at the time of data collection.
* GDP: The Gross Domestic Product at the time of data collection.
* Target: The target variable indicating student outcome (e.g., Dropout, Academic Success).


\pagebreak
### Data Attribute Analysis
The following section explores the nature of the attributes in the dataset.


\pagebreak

### Data Transformations

```{r echo=FALSE}
# Perform necessary transformations

# Fill missing values with the mean for numeric columns and mode for categorical columns

studentSuccessData <- studentSuccessData %>%

  mutate(across(where(is.numeric), ~ ifelse(is.na(.), mean(., na.rm = TRUE), .))) %>%

  mutate(across(where(is.character), ~ ifelse(is.na(.), Mode(.), .)))
 
# Function to calculate mode

Mode <- function(x) {

  ux <- unique(x)

  ux[which.max(tabulate(match(x, ux)))]

}
```



```{r echo=FALSE}

# Convert categorical columns to factor type

studentSuccessData <- studentSuccessData %>%

  mutate(

    Marital.status = factor(
      Marital.status,
      levels = c(1, 2, 3, 4, 5, 6),
      labels = c("Single", "Married", "Widower", "Divorced", "Facto Union", "Legally Separated")),

    Application.mode = factor(
      Application.mode,
      levels = c(1, 2, 5, 7, 10, 15, 16, 17, 18, 26, 27, 39, 42, 43, 44, 51, 53, 57),
      labels = c(
        "1st phase - general contingent",
        "Ordinance No. 612/93",
        "1st phase - special contingent (Azores Island)",
        "Holders of other higher courses",
        "Ordinance No. 854-B/99",
        "International student (bachelor)",
        "1st phase - special contingent (Madeira Island)",
        "2nd phase - general contingent",
        "3rd phase - general contingent",
        "Ordinance No. 533-A/99, item b2) (Different Plan)",
        "Ordinance No. 533-A/99, item b3 (Other Institution)",
        "Over 23 years old",
        "Transfer",
        "Change of course",
        "Technological specialization diploma holders",
        "Change of institution/course",
        "Short cycle diploma holders",
        "Change of institution/course (International)")),

    Application.order = factor(Application.order, ordered=TRUE, levels = c(1, 2, 3, 4, 5, 6), labels = c("1st", "2nd", "3rd", "4th", "5th", "6th")),
    
    Target = factor(Target),

    Course = factor(Course, levels = c(33,171,8014,9003,9070,9085,9119,9130,9147,9238,9254,9500,9556,9670,9773,9853,9991), labels = c(
      "Biofuel Production Technologies",
      "Animation and Multimedia Design",
      "Social Service (evening attendance) ",
      "Agronomy",
      "Communication Design",
      "Veterinary Nursing",
      "Informatics Engineering",
      "Equinculture",
      "Management",
      "Social Service",
      "Tourism",
      "Nursing",
      "Oral Hygiene",
      "Advertising and Marketing Management",
      "Journalism and Communication",
      "Basic Education",
      "Management (evening attendance)")),


    Previous.qualification = factor(
      Previous.qualification,
      levels = c(1, 2, 3, 4, 5, 6, 9, 10, 12, 14, 15, 19, 38, 39, 40, 42, 43),
      labels = c(
        "Secondary education",
        "Higher education - bachelor's degree",
        "Higher education - degree",
        "Higher education - master's",
        "Higher education - doctorate",
        "Frequency of higher education",
        "12th year of schooling - not completed",
        "11th year of schooling - not completed",
        "Other - 11th year of schooling",
        "10th year of schooling",
        "10th year of schooling - not completed",
        "Basic education 3rd cycle (9th/10th/11th year) or equiv.",
        "Basic education 2nd cycle (6th/7th/8th year) or equiv.",
        "Technological specialization course",
        "Higher education - degree (1st cycle)",
        "Professional higher technical course",
        "Higher education - master (2nd cycle)")),

    Nacionality = factor(
      Nacionality,
      levels = c(1, 2, 6, 11, 13, 14, 17, 21, 22, 24, 25, 26, 32, 41, 62, 100, 101, 103, 105, 108, 109),
      labels = c(
        "Portuguese",
        "German",
        "Spanish",
        "Italian",
        "Dutch",
        "English",
        "Lithuanian",
        "Angolan",
        "Cape Verdean",
        "Guinean",
        "Mozambican",
        "Santomean",
        "Turkish",
        "Brazilian",
        "Romanian",
        "Moldova (Republic of)",
        "Mexican",
        "Ukrainian",
        "Russian",
        "Cuban",
        "Colombian")),

    # Causes the error.
#     Mother.s.qualification <- ordered(Mother.s.qualification, levels = c(1, 2, 3, 4, 5, 6, 9, 10, 11, 12, 14, 18, 19, 22, 26, 27, 29, 30, 34, 35, 36, 37, 38, 39, 40, 41, 42, 43, 44), labels = c(
#   "Secondary Education - 12th Year of Schooling or Eq.",
#   "Higher Education - Bachelor's Degree",
#   "Higher Education - Degree",
#   "Higher Education - Master's",
#   "Higher Education - Doctorate",
#   "Frequency of Higher Education",
#   "12th Year of Schooling - Not Completed",
#   "11th Year of Schooling - Not Completed",
#   "7th Year (Old)",
#   "Other - 11th Year of Schooling",
#   "10th Year of Schooling",
#   "General commerce course",
#   "Basic Education 3rd Cycle (9th/10th/11th Year) or Equiv.",
#   "Technical-professional course",
#   "7th year of schooling",
#   "2nd cycle of the general high school course",
#   "9th Year of Schooling - Not Completed",
#   "8th year of schooling",
#   "Unknown",
#   "Can't read or write",
#   "Can read without having a 4th year of schooling",
#   "Basic education 1st cycle (4th/5th year) or equiv.",
#   "Basic Education 2nd Cycle (6th/7th/8th Year) or Equiv.",
#   "Technological specialization course",
#   "Higher education - degree (1st cycle)",
#   "Specialized higher studies course",
#   "Professional higher technical course",
#   "Higher Education - Master (2nd cycle)",
#   "Higher Education - Doctorate (3rd cycle)"
# )),
# 
#  Father.s.qualification <- ordered(Father.s.qualification, levels = c(1, 2, 3, 4, 5, 6, 9, 10, 11, 12, 13, 14, 18, 19, 20, 22, 25, 26, 27, 29, 30, 31, 33, 34, 35, 36, 37, 38, 39, 40, 41, 42, 43, 44), labels = c(
#   "Secondary Education - 12th Year of Schooling or Eq.",
#   "Higher Education - Bachelor's Degree",
#   "Higher Education - Degree",
#   "Higher Education - Master's",
#   "Higher Education - Doctorate",
#   "Frequency of Higher Education",
#   "12th Year of Schooling - Not Completed",
#   "11th Year of Schooling - Not Completed",
#   "7th Year (Old)",
#   "Other - 11th Year of Schooling",
#   "2nd year complementary high school course",
#   "10th Year of Schooling",
#   "General commerce course",
#   "Basic Education 3rd Cycle (9th/10th/11th Year) or Equiv.",
#   "Complementary High School Course",
#   "Technical-professional course",
#   "Complementary High School Course - not concluded",
#   "7th year of schooling",
#   "2nd cycle of the general high school course",
#   "9th Year of Schooling - Not Completed",
#   "8th year of schooling",
#   "General Course of Administration and Commerce",
#   "Supplementary Accounting and Administration",
#   "Unknown",
#   "Can't read or write",
#   "Can read without having a 4th year of schooling",
#   "Basic education 1st cycle (4th/5th year) or equiv.",
#   "Basic Education 2nd Cycle (6th/7th/8th Year) or Equiv.",
#   "Technological specialization course",
#   "Higher education - degree (1st cycle)",
#   "Specialized higher studies course",
#   "Professional higher technical course",
#   "Higher Education - Master (2nd cycle)",
#   "Higher Education - Doctorate (3rd cycle)"
# )),
# 
# Mother.s.occupation <- ordered(Mother.s.occupation, levels = c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 90, 99, 122, 123, 125, 131, 132, 134, 141, 143, 144, 151, 152, 153, 171, 173, 175, 191, 192, 193, 194), labels = c(
#   "Student",
#   "Representatives of the Legislative Power and Executive Bodies, Directors, Directors and Executive Managers",
#   "Specialists in Intellectual and Scientific Activities",
#   "Intermediate Level Technicians and Professions",
#   "Administrative staff",
#   "Personal Services, Security and Safety Workers and Sellers",
#   "Farmers and Skilled Workers in Agriculture, Fisheries and Forestry",
#   "Skilled Workers in Industry, Construction and Craftsmen",
#   "Installation and Machine Operators and Assembly Workers",
#   "Unskilled Workers",
#   "Armed Forces Professions",
#   "Other Situation",
#   "(blank)",
#   "Health professionals",
#   "Teachers",
#   "Specialists in information and communication technologies (ICT)",
#   "Intermediate level science and engineering technicians and professions",
#   "Technicians and professionals, of intermediate level of health",
#   "Intermediate level technicians from legal, social, sports, cultural and similar services",
#   "Office workers, secretaries in general and data processing operators",
#   "Data, accounting, statistical, financial services and registry-related operators",
#   "Other administrative support staff",
#   "Personal service workers",
#   "Sellers",
#   "Personal care workers and the like",
#   "Skilled construction workers and the like, except electricians",
#   "Skilled workers in printing, precision instrument manufacturing, jewelers, artisans and the like",
#   "Workers in food processing, woodworking, clothing and other industries and crafts",
#   "Cleaning workers",
#   "Unskilled workers in agriculture, animal production, fisheries and forestry",
#   "Unskilled workers in extractive industry, construction, manufacturing and transport",
#   "Meal preparation assistants"
# )),
# 
#  Father.s.occupation <- factor(Father.s.occupation,
# levels = c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 90, 99, 101, 102, 103, 112, 114, 121, 122, 123, 124, 131, 132, 134, 135, 141, 143, 144, 151, 152, 153, 154, 161, 163, 171, 172, 174, 175, 181, 182, 183, 192, 193, 194, 195), labels = c(
#   "Student",
#   "Representatives of the Legislative Power and Executive Bodies, Directors, Directors and Executive Managers",
#   "Specialists in Intellectual and Scientific Activities",
#   "Intermediate Level Technicians and Professions",
#   "Administrative staff",
#   "Personal Services, Security and Safety Workers and Sellers",
#   "Farmers and Skilled Workers in Agriculture, Fisheries and Forestry",
#   "Skilled Workers in Industry, Construction and Craftsmen",
#   "Installation and Machine Operators and Assembly Workers",
#   "Unskilled Workers",
#   "Armed Forces Professions",
#   "Other Situation",
#   "(blank)",
#   "Armed Forces Officers",
#   "Armed Forces Sergeants",
#   "Other Armed Forces personnel",
#   "Directors of administrative and commercial services",
#   "Hotel, catering, trade and other services directors",
#   "Specialists in the physical sciences, mathematics, engineering and related techniques",
#   "Health professionals",
#   "Teachers",
#   "Specialists in finance, accounting, administrative organization, public and commercial relations",
#   "Intermediate level science and engineering technicians and professions",
#   "Technicians and professionals, of intermediate level of health",
#   "Intermediate level technicians from legal, social, sports, cultural and similar services",
#   "Information and communication technology technicians",
#   "Office workers, secretaries in general and data processing operators",
#   "Data, accounting, statistical, financial services and registry-related operators",
#   "Other administrative support staff",
#   "Personal service workers",
#   "Sellers",
#   "Personal care workers and the like",
#   "Protection and security services personnel",
#   "Market-oriented farmers and skilled agricultural and animal production workers",
#   "Farmers, livestock keepers, fishermen, hunters and gatherers, subsistence",
#   "Skilled construction workers and the like, except electricians",
#   "Skilled workers in metallurgy, metalworking and similar",
#   "Skilled workers in electricity and electronics",
#   "Workers in food processing, woodworking, clothing and other industries and crafts",
#   "Fixed plant and machine operators",
#   "Assembly workers",
#   "Vehicle drivers and mobile equipment operators",
#   "Unskilled workers in agriculture, animal production, fisheries and forestry",
#   "Unskilled workers in extractive industry, construction, manufacturing and transport",
#   "Meal preparation assistants",
#   "Street vendors (except food) and street service providers"
# )),
# Displaced <- factor(Displaced, levels = c(1, 0), labels = c("yes", "no")),
# Educational.special.needs <- factor(Educational.special.needs, levels = c(1, 0), labels = c("yes", "no")),
# Debtor <- factor(Debtor, levels = c(1, 0), labels = c("yes", "no")),
# Tuition.fees.up.to.date <- factor(Tuition.fees.up.to.date, levels = c(1, 0), labels = c("yes", "no")),
# Gender <- factor(Gender, levels = c(1, 0), labels = c("yes", "no")),
# Scholarship.holder <- factor(Scholarship.holder, levels = c(1, 0), labels = c("yes", "no")),
# International <- factor(International, levels = c(1, 0), labels = c("yes", "no")),
    # Daytime.evening.attendance. <- factor(Daytime.evening.attendance., levels = c(1, 0), labels = c("daytime", "evening"))
)
```


```{r}
#validate structure after attribute transformation
str(studentSuccessData)
```


#### Data with factors

```{r}
summary(studentSuccessData)
```


#### Continuous Attributes

These attributes are continuous variables:
["Unemployment.rate", "Inflation.rate", "GDP"]                            


\pagebreak
### Exploratory Data Analysis

Divide the data into a reserved test set (30%) and a train set (70%).

```{r}
set.seed(711)
ind <- sample(c(TRUE, FALSE), size=nrow(studentSuccessData), replace=TRUE, prob=c(0.7, 0.3))
train <- studentSuccessData[ind, ]
test <- studentSuccessData[!ind, ]
```

```{r}
colnames(train)
```


```{r}
 
columns <- colnames(train)
# Loop through each column and create a plot
for (col in columns) {
  # Check if the column is numeric or categorical
  if (is.numeric(train[[col]])) {
    # Create a histogram for numeric columns
    p <- ggplot(train, aes_string(x = col)) +
      geom_histogram(binwidth = 1, fill = "blue", color = "black") +
      ggtitle(paste("Histogram of", col)) +
      theme_minimal()
  } else {
    # Create a bar plot for categorical columns
    p <- ggplot(train, aes_string(x = col)) +
      geom_bar(fill = "blue", color = "black") +
      ggtitle(paste("Bar Plot of", col)) +
      theme_minimal()
  }
  # Print the plot
  print(p)
}
```


Here is an example small multiple.

```{r}
#base_graph <- ggplot(train, aes(Target)) + geom_bar()

#base_graph + facet_grid(cols=vars(Marital.status), rows=vars(Nacionality))
```

*** Bi-Variate EDA

```{r}

# Plot for Age and Target

ggplot(train, aes(x = Age.at.enrollment, y = Target)) +
  geom_point() +
  geom_smooth(method = "lm", col = "blue") +
  labs(title = "Scatter Plot of Age.at.enrollment  vs Target",
       x = "Age.at.enrollment ",
       y = "Target") +
  theme_minimal()

```


### 2. Mother's Qualification and Target
```{r}
# Bar plot for Mother's Qualification and Target
ggplot(train, aes(x = factor(Mother.s.qualification), fill = factor(Target))) +
  geom_bar(position = "dodge") +
  labs(title = "Bar Plot of Mother's Qualification vs Target",
       x = "Mother's Qualification",
       fill = "Target") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set3") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### 3. Father's Qualification and Target
```{r}
# Bar plot for Father's Qualification and Target
ggplot(train, aes(x = factor(Father.s.qualification), fill = factor(Target))) +
  geom_bar(position = "dodge") +
  labs(title = "Bar Plot of Father's Qualification vs Target",
       x = "Father's Qualification",
       fill = "Target") +
  theme_minimal() +
  scale_fill_manual(values = c("#DC143C", "#1f77b4", "#2ca02c"))  

```

### 4. Mother's Occupation, Father's Occupation, and Target
```{r}
# Bar plot for Father's Occupation and Target
ggplot(train, aes(x = factor(Father.s.occupation), fill = factor(Target))) +
  geom_bar(position = "dodge") +
  labs(title = "Bar Plot of Father's Occupation vs Target",
       x = "Father's Occupation",
       fill = "Target") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set3") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Bar plot for Mother's Occupation and Target
ggplot(train, aes(x = factor(Mother.s.occupation), fill = factor(Target))) +
  geom_bar(position = "dodge") +
  labs(title = "Bar Plot of Mother's Occupation vs Target",
       x = "Mother's Occupation",
       fill = "Target") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set3") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

### 5. Previous Qualification and Target
```{r}
# Bar plot for Previous Qualification and Target
ggplot(train, aes(x = factor(Previous.qualification), fill = factor(Target))) +
  geom_bar(position = "dodge")  +
  labs(title = "Bar Plot of Previous Qualification vs Target",
       x = "Previous Qualification",
       fill = "Target") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set3") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```




\pagebreak
## Models

### Model 1 Association Rule Mining



#### Step 1
```{r}
# For association rule mining, we need to convert the data into transactions:
library(arules)

train.arm <- train

# Identify numeric columns
numeric_columns <- sapply(train, is.numeric)

# Discretize numeric columns
for (col in names(train)[numeric_columns]) {
  train.arm[[col]] <- discretize(train[[col]], method = "interval", breaks = 5)
}

transaction_id  <- as.character(train.arm[["transid"]])
train.arm$transid <- NULL
transactions <- as(train.arm, "transactions")

```


#### Step 2
```{r}
# Generate association rules
rules_dropout <- apriori(transactions, parameter = list(supp = 0.01, conf = 0.5, target = "rules"),
                 appearance = list(rhs = c("Target=Dropout" ), default = "lhs") , control = list(verbose=F))
rules_graduate <- apriori(transactions, parameter = list(supp = 0.01, conf = 0.5, target = "rules"),
                 appearance = list(rhs = c("Target=Graduate" ), default = "lhs") , control = list(verbose=F))
```


#### Step 3
```{r}

library(arulesViz)

# Inspect the top rules of Dropout by lift
rules_dropout_sorted <- sort(rules_dropout,decreasing = TRUE, by = "lift")
inspect(rules_dropout_sorted[1:10])

# Inspect the top rules of Dropout by confidence
rules_dropout_sorted <- sort(rules_dropout_sorted,decreasing = TRUE, by = "confidence")
inspect(rules_dropout_sorted[1:10])

#summarize the rules
summary(rules_dropout_sorted)

#visualize
plot(rules_dropout_sorted[1:10], method="graph", engine = "interactive")

```
```{r}

# Inspect Top rules of graduate

rules_graduate_sorted <- sort(rules_graduate ,decreasing = TRUE, by = "lift")
inspect(rules_graduate_sorted[1:10])
rules_graduate_sorted <- sort(rules_graduate_sorted,decreasing = TRUE, by = "confidence")
inspect(rules_graduate_sorted[1:10])
summary(rules_graduate_sorted)
plot(rules_graduate_sorted[1:10], method="graph", engine = "interactive")
```







Here are the clusters grouped by Target.  
```{r}
# ggplot(data=emdata, aes(x=Target, fill=cluster)) +
#     geom_bar(stat="count") +
#     labs(title = "K = ?") +
#     theme(plot.title = element_text(hjust=0.5),  text=element_text(size=15))
```


## Decision Tree Model  ## rpart Model  ## rpart Model performance

### Model 3 Decision Tree Classification 

```{r}

### For decision tree classification, we'll use the `rpart` package:

# We need to split the dataset, but since we already partitioned the dataset into train and test, we will directly use that for training our model.

# Train the decision tree model
tree_model <- rpart(Target ~ ., data = train, method = "class")

# Prune the tree to simplify it
pruned_tree <- prune(tree_model, cp = tree_model$cptable[which.min(tree_model$cptable[,"xerror"]),"CP"])

# Plot the pruned tree with improved readability
par(mar = c(5, 4, 4, 2) + 0.1)  # Adjust the margins as needed

rpart.plot(pruned_tree, 
           type = 2, 
           extra = 104, 
           fallen.leaves = TRUE, 
           cex = 0.8,  # Adjust text size
           box.palette = "RdYlGn", 
           shadow.col = "gray", 
           nn = TRUE,
           faclen = 0,  # Full names
           varlen = 0,  # Full variable names
           compress = TRUE)  # Compress the tree



text(tree_model, use.n = TRUE)

#capture the values as reference to compare in the confusion matrix
test.target <- test$Target


#drop the orginal target value
test %>% select(-Target)


#create the predictions
predictions <- predict(tree_model, newdata=test, type = "class")


# Confusion matrix
confusionMatrix(predictions, test.target)

```



## Additional Models

\pagebreak
# Conclusions



\pagebreak
# Limitations of the Analysis and/or Available Data




\pagebreak
# Bibliography

Realinho,Valentim, Vieira Martins,Mónica, Machado,Jorge, and Baptista,Luís. (2021). Predict Students' Dropout and Academic Success. UCI Machine Learning Repository. https://doi.org/10.24432/C5MC89.

